DRAFT\\Satisfiability Modulo Theories Competition (SMT-COMP) 2014: Rules and Procedures

### 2014-05-19 rules/rules14.tex
slight edits
Rule revisions

The rest of this document, revised from the previous 
version,\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.}
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from the previous competition rules are the following:
\begin{itemize}
\item The competition will be run with StarExec (\url{www.starexec.org}). Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [ The demonstration unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.]
\item In the main track, we will measure comparative performance within all divisions to which at least two solvers are submitted. We will use the data to determine both sequential and parallel winners.

\item We are retaining and accenting last year's ``application track" and
  encourage submission of benchmarks for this track and competition against these
  benchmarks.

@@

\header{System description.} %
As part of a submission,
SMT-COMP entrants are encouraged to provide a short (1--2 pages) description of
the system.  This should include a list of all authors of the system,
 their present institutional affiliations, and any appropriate acknowledgements.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool.

@@

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via StarExec and email to the organizers by the end of May~25, 2014 (anywhere on earth).  At that time new entrants will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until the end of June~1, 2014 (anywhere on earth).  We strongly encourage
participants to use this grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using StarExec after the original deadline on May~25.

The versions that are present on the execution service at the conclusion of the
grace period will be the ones used for the competition, and versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the morning of June~2.  See
Section~\ref{sec:timeline} below for a full timeline.

@@

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the weeks leading up to SMT 2014, from about June~2 to July~17.  Results will be
announced at SMT 2014 and in an awards ceremony scheduled by the FLoC Olympic Games coordinator.
  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

@@

\header{Input and Output.}
%
Participating solvers must read a single benchmark script presented as its first argument.
Note that previous competitions presented benchmarks to a solver on its standard input
channel, and that the application track will send command inputs to the solver's standard input.
Accordingly, solvers are advised to support both input mechanisms.

@@

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command after any $\akey{set-option}$ commands described below.
% rb 25/3: added
% \item A (single) \akey{set-option :print-success false} command, sent just
%       after $\akey{set-logic}$. This is to
%       avoid the output ``\texttt{success}'' to interfere with the
%       answer of the solver to the benchmark. 
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item For the application track, there are one or more \akey{check-sat} commands, 
  each preceded by one or more \akey{assert} commands 
  and zero or more \akey{push 1} commands, 
  and followed by zero or more \akey{pop 1} commands.
\item Scripts for the application track will have an initial $\akey{set-option :print-success true}$ command.
\item There is at most one $\akey{set-info}$ command for \texttt{status}.
%%%%%%%%%%%%%%%%%%% TODO: Need to define the set-info command.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared exactly once before any
  use, using $\akey{declare-sort}$
  %, $\akey{define-sort}$,
  %$\akey{declare-fun}$, or $\akey{define-fun}$.  
  or $\akey{declare-fun}$.
  They must be part of the allowed signature expansion for the logic.
  Moreover, all sorts declared with a $\akey{declare-sort}$ command must have zero arity.

\item No other commands besides the ones just mentioned may be used.

@@

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given benchmark, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of CPU time and the same limit on wall clock time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
% TBD- TODO - how to constrain solvers to single processes?
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.  A
timeout scores the same as if the output is ``\texttt{unknown}''.

@@

@@

\medskip
Solvers must respond to each command sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands. There should be no \akey{get-unsat-core} or \akey{get-proof} commands in these benchmarks.

@@

\header{Benchmark availability.} % 
The deadline for new benchmarks is May 15, 2014.
The benchmarks will be made available as soon as possible after that, as they are checked and curated.
The goal for completing the set of benchmarks is June 1.
The set of selected benchmarks will be published when the
competition begins.

@@

The benchmarks are assigned a difficulty. For 2014, the difficulty measure is the
CPU time taken by the best performing solver in the 2013 SMT-EVAL. These times will be posted 
as soon after May 15 as possible.

@@

Then, for each division, the benchmarks are divided into quintiles according to the difficulty measure.
Then a random selection of N competition benchmarks is made as follows:
\begin{itemize}
\item N*40\% are taken from the top quintile, if sufficient benchmarks are available
\item additional benchmarks are taken from the second quintile, to produce N*60\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the third quintile, to produce N*75\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the fourth quintile, to produce N*90\% benchmarks, if sufficient benchmarks are available
\item any remaining needed benchmarks are taken from the bottom quintile.
\end{itemize}
Within each quintile selection,
\begin{itemize}
\item if the industrial benchmarks comprise less than 85\% of the benchmarks in that quintile, 85\% will be chosen from the industrial benchmarks (if a sufficient number are available), and the others from the population of random and crafted benchmarks;
\item if the industrial benchmarks comprise at least 85\% of the benchmarks in that quintile, the benchmarks are chosen at random from the population of industrial, random and crafted benchmarks in that quintile.
\end{itemize}

Since the organizers at this point are unsure how long a set of benchmarks may take (which will depend also on the number of solvers submitted), the competition will be run in {\em heats}. For each division, some number $N$ of benchmarks will be selected as described and randomly divided into a number of heats. The heats will be run in order. If the competition ends before all heats have completed, the remaining heats will be ignored. All benchmarks will be chosen at the beginning (that is, without reference to the results of any given heat).

@@

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on
the first day the exchange is open on or after June 2
will be added to the other seeding values.  This helps provide transparency,
by guaranteeing that the organizers cannot manipulate the seed in
favor of or against any particular submitted solver.  Benchmarks will also be slightly
scrambled before the competition, using a simple benchmark scrambler
seeded with the same seed as the benchmark selector.  Both the
scrambler and benchmark selector will be publicly available before the
competition.  Naturally, solvers must not rely on previously
determined identifying syntactic characteristics of competition
benchmarks in testing satisfiability (violation of this is considered
cheating).

@@

\header{Benchmark availability.} 
A first release of the application track
benchmarks will be made available May 15, 2014.
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues. 
The final selection that will be used for the competition will be posted when the competition begins.

@@

\header{Competition scoring.}
Benchmarks' scores are summed component-wise to form a solver's total
score for that division of the competition.
Total scores are compared lexicographically---a score $\langle e,n,m\rangle$ is better than 
$\langle e',n',m'\rangle$ iff $e < e'$ or ($e = e'$ and $n > n'$) or ($e = e'$ and $n = n'$ and $m < m'$).
That is, fewer errors takes precedence over more correct solutions, which takes precedence over less time taken.

\subsection{Other recognitions}
The organizers will also recognize the following contributions:
\begin{itemize}
\item {\em Open source}. In addition to recognizing the overall
winner in each division, the top solver that provides its source code as open source will also be
recognized in each division.  
\item {\em Best new entrant}. The best performing entrant from a new solver implementation team, as measured by the average of 
(solver time)/(winning solver time) over the competition benchmarks.
\item The organizers reserve the right to recognize other outstanding contributions that become apparent in the competition results.
\end{itemize}

@@

The FLoC Olympics games medals will be awarded based on competition-wide critera. The organizers are still determining what those criteria should be. Comments are welcome.

%% TBD: DD suggestion: We could have a system similar to that of the Diamond's league. For each problem division with at least three entrants, attribute 4, 2 and 1 points to the first three solvers. The medals will be awarded based on the total number of points.

@@

\nobreak
\vbox{% no page break here, please!
\begin{description}
\item[May 15] Deadline for benchmark contributions.
  First version of the benchmark scrambler, benchmark selector and trace executor made available.
\item[May 25] First versions of solver competitors are due to the organizers.
\item[June 1] Final version of the benchmark scrambler, benchmark selector and trace executor made available.
\item[June 1] Benchmark libraries are frozen; final versions of solver competitors due.
\item[June 1] Final versions of solvers due via StarExec (for
  all tracks), including divisions being entered, system
  descriptions and magic numbers for benchmark scrambling.
\item[June 2] Opening value of NYSE Composite Index used to complete random seed.
\item[July 17] SMT 2014; end of competition.
\end{description}}

@@

\section{Disclaimer}
\begin{itemize}
\item David Cok, David Deharbe and Tjark Weber are the organizers of SMT-COMP 2014. 
David Cok is the lead and is 
responsible for policy and procedure decisions, such as these
rules, with input from the co-organizers. He is not associated 
with any group creating or submitting solvers. He has used solvers
in industrial settings and is keenly interested to know which are the best.

>>>>

The rest of this document, revised from the previous 
version,\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.}
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from the previous competition rules are the following:
\begin{itemize}
\item The competition will be run with StarExec (\url{www.starexec.org}). Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [ The demonstration unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.] We encourage submission of solvers and benchmarks for both tracks.
\item In the main track, we will measure comparative performance within all divisions to which at least two solvers are submitted. We will use the data to determine both sequential and parallel winners.

@@

\header{System description.} %
As part of a submission,
SMT-COMP entrants are encouraged to provide a short (1--2 pages) description of
the system.  This should include a list of all authors of the system,
 their present institutional affiliations, and any appropriate acknowledgements.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool.  System descriptions may be submitted after the solver deadline, but to be useful should be sent to the organizers before the competition ends (July 16).

@@

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via StarExec and email to the organizers by the end of June~1, 2014 (anywhere on earth).  At that time new entrants will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until the end of June~15, 2014 (anywhere on earth).  We strongly encourage
participants to use this grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using StarExec after the initial deadline.

The versions that are present on the execution service at the conclusion of the
grace period will be the ones used for the competition, and versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the day after the final solver deadline.  See
Section~\ref{sec:timeline} below for a full timeline.

@@

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the weeks leading up to SMT 2014, from about June~16 to July~17.  Results will be
announced at SMT 2014 and in an awards ceremony scheduled by the FLoC Olympic Games coordinator.
  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

@@

\header{Input and Output.}
%
Participating solvers must read a single benchmark script presented as its first argument.
Note that previous competitions presented benchmarks to a solver on its standard input
channel, and that the application track will send command inputs to the solver's standard input.
Accordingly, solvers are advised to support both input mechanisms. The organizers have easily adapted solvers that use only the standard input to Star-Exec with a simple wrapper script.

@@

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command after any $\akey{set-option}$ commands described below.
\item In the main track, there may be a (single) \akey{set-option :print-success ...} command. Note that {\tt success} outputs are ignored by the post-processor used by the competition.
\item Scripts for the application track will have an initial $\akey{set-option :print-success true}$ command.
\item The script next optionally contains (any number of) $\akey{set-info}$ commands, which may be ignored by the solver.
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item For the application track, there are one or more \akey{check-sat} commands, 
  each preceded by one or more \akey{assert} commands 
  and zero or more \akey{push 1} commands, 
  and followed by zero or more \akey{pop 1} commands.

%%%%%%%%%%%%%%%%%%% TODO: Need to define the set-info command.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared exactly once before any
  use, using $\akey{declare-sort}$
  %, $\akey{define-sort}$,
  %$\akey{declare-fun}$, or $\akey{define-fun}$.  
  or $\akey{declare-fun}$.
  They must be part of the allowed signature expansion for the logic.
  Moreover, all sorts declared with a $\akey{declare-sort}$ command must have zero arity.

\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.

@@

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given benchmark, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of CPU time and the same limit on wall clock time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.  A
timeout scores the same as if the output is ``\texttt{unknown}''.

@@

The solver is executed with a temporary directory as the current working directory.
Any generated files should be produced there, not, for example, in the system tmp area.
The temporary directory and its contents are deleted after the job is complete.

@@

\medskip
Solvers must respond to each command sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands. There should be no \akey{get-unsat-core}, \akey{get-proof}, \akey{get-model}, \akey{get-assignments}, or \akey{get-values}commands in these benchmarks.

@@

\header{Benchmark availability.} % 
The deadline for new benchmarks is May 15, 2014.
The benchmarks will be made available as soon as possible after that, as they are checked and curated.
The goal for completing the set of benchmarks is June 1.
The set of benchmarks selected for the competition will be published when the
competition begins.

@@

The benchmarks are assigned a difficulty. For 2014, the difficulty measure is the
CPU time taken by the best performing solver in the 2013 SMT-EVAL. 
 For new benchmarks, we will determine an approximate difficulty
using trial runs.
The difficulty values will be posted 
as soon after May 15 as possible.
The difficulty values are best estimates by the organizers and may not always completely correlate with actual difficulty.

@@

Then, for each division, the benchmarks are divided into quintiles according to the difficulty measure.
Next a random selection of N competition benchmarks is made as follows:
\begin{itemize}
\item N*40\% are randomly selected from the top quintile, if sufficient benchmarks are available
\item additional benchmarks are randomly selected from the second quintile, to produce N*60\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are randomly selected from the third quintile, to produce N*75\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are randomly selected from the fourth quintile, to produce N*90\% benchmarks, if sufficient benchmarks are available
\item any remaining needed benchmarks are randomly selected from the bottom quintile.
\end{itemize}
Within each quintile selection,
\begin{itemize}
\item if the industrial benchmarks comprise less than 85\% of the benchmarks in that quintile, 85\% will be chosen from the industrial benchmarks (if a sufficient number are available), and the others from the population of random and crafted benchmarks;
\item if the industrial benchmarks comprise at least 85\% of the benchmarks in that quintile, the benchmarks are chosen at random from the population of industrial, random and crafted benchmarks in that quintile.
\end{itemize}

Since the organizers at this point are unsure how long a set of benchmarks may take (which will depend also on the number of solvers submitted), the competition will be run in {\em heats}. For each division, some number $N$ of benchmarks will be selected as described and randomly divided into a number of (possibly unequal-sized) heats. The heats will be run in order. If the competition ends before all heats have completed, the remaining heats will be ignored. All benchmarks will be chosen at the beginning (that is, without reference to the results of any given heat).

@@

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on
the first day the exchange is open on or after the date specified in the timeline
will be added to the other seeding values.  This helps provide transparency,
by guaranteeing that the organizers cannot manipulate the seed in
favor of or against any particular submitted solver.  Benchmarks will also be slightly
scrambled before the competition, using a simple benchmark scrambler
seeded with the same seed as the benchmark selector.  Both the
scrambler and benchmark selector will be publicly available before the
competition.  Naturally, solvers must not rely on previously
determined identifying syntactic characteristics of competition
benchmarks in testing satisfiability (violation of this is considered
cheating).

@@

\header{Benchmark availability.} 
A first release of the application track
benchmarks will be made available as soon after the benchmark deadline as the organizers can arrange.
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues. 
The final selection that will be used for the competition will be posted when the competition begins.

@@

\header{Competition scoring.}
\label{scoring}
Benchmarks' scores are summed component-wise to form a solver's total
score for that division of the competition.
Total scores are compared lexicographically---a score $\langle e,n,m\rangle$ is better than 
$\langle e',n',m'\rangle$ iff $e < e'$ or ($e = e'$ and $n > n'$) or ($e = e'$ and $n = n'$ and $m < m'$).
That is, fewer errors takes precedence over more correct solutions, which takes precedence over less time taken.

\subsection{Other recognitions}
The organizers will also recognize the following contributions:
\begin{itemize}
\item {\em Open source}. In addition to recognizing the overall
winner in each division, the top solver that provides its source code as open source will also be
recognized in each division.  
\item {\em Best new entrant}. The best performing entrant from a new solver implementation team, as measured by the Olympic Games metric defined below.
\item {\em Breadth of logics}. Tools that cover the most theories and logics.
\item {\em Benchmarks}. Contributors of new benchmarks.
\item The organizers reserve the right to recognize other outstanding contributions that become apparent in the competition results.
\end{itemize}

@@

The FLoC Olympics games medals will be awarded based on competition-wide critera.

We define a competition-wide metric M as follows. Each solver competing in a division receives a number of points for its performance in that division: $(n/N)^2$, where $n$ is the number of benchmarks solved correctly and $N$ is the number of benchmarks used in that division. A solver's score for the whole competition is the sum of these values over all the 
divisions in which the solver participates.

The Olympic Games medals will be awarded as follows.
\begin{itemize}
\item The gold medal we will be awarded to the solver with the best competition-wide value of M.
\item The silver medal will be awarded to the solver that wins the (sequential) QF\_BV division according to the usual $\langle e',n',m'\rangle$ metric of subsection \ref{scoring}.
\item The bronze medal will be awarded to the solver with the second-best  competition-wide value of M.
\end{itemize}

{\em Rationale.} The metric M purposely emphasizes breadth of solver participation - a solver participating in many logics need not be the best in any one of them. The use of the square in the metric limits that somewhat - a solver still needs to do reasonably well compared to winners to be able to catch up by breadth of participation. The non-linear metric also gives added weight to completing close to all benchmarks in a division. The metric is also quite simple and the metric for a solver is independent of the performance of other solvers.

@@

\nobreak
\vbox{% no page break here, please!
\begin{description}
\item[May 15] Deadline for benchmark contributions.
\item[June 1] First versions of solver competitors are due to the organizers, including divisions being entered.
  Versions of the benchmark scrambler, benchmark selector and trace executor made available.
\item[June 8] Final version of the benchmark scrambler, benchmark selector and trace executor made available.
\item[June 8] Benchmark libraries are frozen.
\item[June 15] Final versions of solvers due via StarExec (for
  all tracks), including divisions being entered, system
  descriptions and magic numbers for benchmark scrambling.
\item[June 16] Opening value of NYSE Composite Index used to complete random seed.
\item[July 17] SMT 2014; end of competition.
\end{description}}

@@

\section{Disclaimer}
\begin{itemize}
\item David Cok, David Deharbe and Tjark Weber are the organizers of SMT-COMP 2014. 
David Cok is the lead and is 
responsible for policy and procedure decisions, such as these
rules, with input from the co-organizers. He is not associated 
with any group creating or submitting solvers. He has used solvers
in industrial settings and is keenly interested to know which are the best for various applications.
